---
title: "Summary Statistics for Casco Bay OA Data 2015-2018"
author: Curtis C. Bohlen
output:
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

# Load Libraries
```{r load_libraries}
library(tidyverse)
library(readr)

library(CBEPgraphics)

load_cbep_fonts()
```

# Load Data
## Establish Folder References
```{r folder_refs}
sibfldnm <- 'Derived_Data'
parent   <- dirname(getwd())
sibling  <- file.path(parent,sibfldnm)

fn    <- 'CascoBayOAData.csv'
fpath <- file.path(sibling,fn)
```


## Load The Data
The following loads existing data, including a "Temperature Adjusted" pCO~2~ value based on Takehashi et al. 2002. It then collapses that data to daily summaries. 

```{r load_data} 
all_data <- read_csv(fpath,
                     col_types = cols(dd = col_integer(), 
                                      doy = col_integer(),
                                      hh = col_integer(),
                                      mm = col_integer(),
                                      yyyy = col_integer())) %>%
  select(c(13, 1:4, 14, 5:6, 8, 7 ,16, 9:12))

daily_data <- all_data %>%
  select(-hh, -yyyy, -mm, -dd, -doy) %>%         # Will recalculate these 
  mutate(the_date = as.Date(datetime)) %>%
  select(-datetime) %>%
  group_by(the_date) %>%
  summarise_at(c("temp", "sal", "co2", "co2_corr", "do", "ph", 'omega_a'),
               c(m    = function(x) median(x, na.rm=TRUE),
                 r    = function(x) {suppressWarnings(max(x, na.rm=TRUE) -
                                                        min(x, na.rm=TRUE))},
                iqr  = function(x) IQR(x, na.rm=TRUE),
                p80r = function(x) {as.numeric(quantile(x, 0.90, na.rm=TRUE) -
                       quantile(x, 0.10, na.rm=TRUE))})) %>%
  mutate(yyyy = as.numeric(format(the_date, format = '%Y')),
         mm   = as.numeric(format(the_date, format = '%m')),
         dd   = as.numeric(format(the_date, format = '%d')),
         doy  = as.numeric(format(the_date, format = '%j')),
         Month = factor(mm, levels=1:12, labels = month.abb)
         )
```

# Overall Summary Statistics
This is legacy code.  It would be easier today to develop this directly in the tidyverse.
```{r summary_stats}
the.mins     <- sapply(all_data[7:15], min, na.rm=TRUE)
the.medians  <- sapply(all_data[7:15], median, na.rm=TRUE)
the.means    <- sapply(all_data[7:15], mean, na.rm=TRUE)
the.maxes    <- sapply(all_data[7:15], max, na.rm=TRUE)
the.SDs  <-   sapply(all_data[7:15], sd, na.rm=TRUE)
the.samplesizes <-  sapply(all_data[7:15], function(x) sum(! is.na(x)) )
result   <-  cbind(the.mins, the.medians, the.means, the.maxes, the.SDs, the.samplesizes)
colnames(result) <- c('Minimum', 'Median', 'Mean', 'Maximum', 'Std. Deviation', 'Observations')
rownames(result) <- c('Temperature',
                      'Salinity',
                      'DO',
                      'pCO2',
                      'pCO2_corr',
                      'pH',
                      'Omega Aragonite',
                      'Omega Calcite',
                      'TA'
                      )
knitr::kable(result, digits = c(1,1,2,1,3.0))
write.csv(result, 'summarystats.csv')
```

```{r cleanup, echo=FALSE}
rm(the.mins, the.means, the.medians, the.maxes, the.SDs, the.samplesizes, result)
```

# Omega Aragonite Observations and Percentage Below Levels of Concern
```{r levels_of_concern_1}
below1.5 <- sum(all_data$omega_a<1.5, na.rm=TRUE)
below1.0 <- sum(all_data$omega_a<1.0, na.rm=TRUE)
TotObs   <- sum(! is.na(all_data$omega_a))
pctbelow1.5 <- below1.5/TotObs
pctbelow1.0 <- below1.0/TotObs

res <- unlist( list(`Count Below 1.0` = below1.0, `Count Below 1.5` = below1.5,
      `Observations` = TotObs,
      `Percent Below 1.0` = pctbelow1.0,
      `Percent Below 1.5` =pctbelow1.5))
rm(below1.0, below1.5, TotObs, pctbelow1.0, pctbelow1.5)
knitr::kable(t(res), digits = c(0,0,0,3,3))
```

# Daily Omega Aragonite (medians) Observations and and Percentage Below Levels of Concern
```{r daily_levels_of_concern}
below1.5 <- sum(daily_data$omega_a_m<1.5, na.rm=TRUE)
below1.0 <- sum(daily_data$omega_a_m<1.0, na.rm=TRUE)
TotObs   <- sum(! is.na(daily_data$omega_a_m))
pctbelow1.5 <- below1.5/TotObs
pctbelow1.0 <- below1.0/TotObs

res <- unlist(list(`Count Below 1.0` = below1.0, `Count Below 1.5` = below1.5,
      `Observations` = TotObs,
      `Percent Below 1.0` = pctbelow1.0,
      `Percent Below 1.5` =pctbelow1.5))
rm(below1.0, below1.5, TotObs, pctbelow1.0, pctbelow1.5)
knitr::kable(t(res), digits = c(0,0,0,3,3))
```

# Monthly Summary Statistics
This is means **across** years.  This is NOT the same as an estimated monthly average, adjusted for year to year variation, imbalances in time of day data was collected, etc.  For that, we would need to estimate marginal means from a GAMM. We do not pursue that idea in this notebook.  
```{r monthly_stats}
all_data %>%
  select(datetime, yyyy, mm, do, co2, co2_corr, ph, omega_a) %>%
  mutate(Month  = factor(mm, labels = month.abb)) %>%
  select(-mm) %>%
  group_by(Month) %>%
  summarize_at(vars(do, co2, co2_corr, ph, omega_a),
               c(avg = ~mean(., na.rm=TRUE)))

```


```{r monthly_stats}
test <- all_data %>%
  select(datetime, yyyy, mm, do, co2, co2_corr, ph, omega_a) %>%
  mutate(Month  = factor(mm, labels = month.abb)) %>%
  select(-mm) %>%
  pivot_longer(do:omega_a, names_to= 'parameter', values_to = 'value') %>%
  group_by(Month, parameter) %>%
  summarise(avg = round(mean(value, na.rm=TRUE),2),
            median = round(median(value, na.rm=TRUE),2),
            sd  = round(sd(value, na.rm=TRUE),3),
            count = sum(! is.na(value)))
```
That contains the info we want, but not quite in the format we would like for a compact table.

What we want is a table that has MONTHS across the top, and statistics within parameters down the side.

So, we pivot longer again, then pivot wider in adifferent configuration.

```{r}
test2 <- test %>% ungroup() %>%
  #select(-Month, -parameter) %>%
  pivot_longer( cols = c('avg', 'median', 'sd', 'count'),
               names_to = 'label')
```

```{r}
test3 <- test2 %>%
  pivot_wider(id_cols = c(parameter, label), names_from=Month) 
test3
write_csv(test3, 'Monthly_summaries.csv')
```





